Lesson 4
========================================================
```{r, include=FALSE}
setwd('C:/Users/Jeff/udacity/Data_Analysis_with_R')
```

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
pf <- read.csv('pseudo_facebook.tsv', sep='\t')

qplot(data=pf, x=age, y=friend_count)
```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
ggplot(pf, aes(x=age, y=friend_count)) + geom_point() + xlim(13,90)

summary(pf$age)
```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(pf, aes(x=age, y=friend_count)) + geom_jitter(alpha=1/20) + xlim(13,90)
```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r Coord_trans()}
ggplot(pf, aes(x=age, y=friend_count)) + xlab('Age') + ylab('Friend Count') + 
  geom_point(alpha=1/20, position = position_jitter(h=0)) + 
  xlim(13,90) + coord_trans(y='sqrt')
```


#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
ggplot(pf, aes(x=age, y=friendships_initiated)) + xlab('Age') + ylab('Friendships Initiated') + 
  geom_point(alpha=1/20, position = position_jitter(h=0)) + xlim(13,90) + coord_trans(y='sqrt')
```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
library(dplyr)

# method 1
age_groups <- group_by(pf, age)
pf.fc_by_age <- summarise(age_groups, 
                          friend_count_mean = mean(friend_count), 
                          friend_count_median = median(friend_count), 
                          n=n()) 
pf.fc_by_age <- arrange(pf.fc_by_age, age)

# method 2
pf.fc_by_age <- pf %>% group_by(age) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>%
  arrange(age)

head(pf.fc_by_age)
```

Create your plot!

```{r Conditional Means Plot}
# Plot mean friend count vs. age using a line graph.
# Be sure you use the correct variable names
# and the correct data frame. You should be working
# with the new data frame created from the dplyr
# functions. The data frame is called 'pf.fc_by_age'.
ggplot(pf.fc_by_age, aes(x=age, y=friend_count_mean)) + geom_line()
```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(pf, aes(x=age, y=friend_count)) + xlab('Age') + ylab('Friend Count') + 
  geom_point(alpha=1/20, position = position_jitter(h=0), color='orange') + 
  xlim(13,90) + 
  coord_trans(y='sqrt') + 
  geom_line(stat='summary', fun.y=mean) +
  geom_line(stat ='summary', fun.y=quantile, fun.args=list(probs=0.1), linetype=2, color='blue') + 
  geom_line(stat ='summary', fun.y=quantile, fun.args=list(probs=0.9), linetype=2, color='blue') +
  geom_line(stat ='summary', fun.y=median, color='blue')

ggplot(pf, aes(x=age, y=friend_count)) + xlab('Age') + ylab('Friend Count') + 
  geom_point(alpha=1/20, position = position_jitter(h=0), color='orange') + 
  xlim(13,90) + 
  coord_cartesian(xlim = c(13,70), ylim = c(0,1000)) + 
  geom_line(stat='summary', fun.y=mean) +
  geom_line(stat ='summary', fun.y=quantile, fun.args=list(probs=0.1), linetype=2, color='blue') + 
  geom_line(stat ='summary', fun.y=quantile, fun.args=list(probs=0.9), linetype=2, color='blue') +
  geom_line(stat ='summary', fun.y=median, color='blue')
```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
# rule of thumb:
#   |correlation| > 0.3 is meaningful but small
#   |correlation| > 0.5 is moderate
#   |correlation| > 0.7 is pretty large
cor.test(pf$age, pf$friend_count, method='pearson')

# equivalent code
with(pf, cor.test(age, friend_count, method='pearson'))
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf,age <= 70), cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}
ggplot(pf, aes(x=www_likes_received, y=likes_received)) + geom_point(alpha=0.1, position=position_jitter(h=0)) + scale_x_log10() + scale_y_log10()
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
ggplot(pf, aes(x=www_likes_received, y=likes_received)) + 
  geom_point(alpha=0.1, position=position_jitter(h=0)) + 
  xlim(0, quantile(pf$www_likes_received, 0.95)) +
  ylim(0, quantile(pf$likes_received, 0.95)) +
  geom_smooth(method='lm', color='red')
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
cor.test(pf$www_likes_received, pf$likes_received)
```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
library(alr3)
data(Mitchell)
```

Create your plot!

```{r Temp vs Month}
ggplot(Mitchell, aes(x=Month, y=Temp)) + geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
with(Mitchell, cor.test(Temp,Month))
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot(Mitchell, aes(x=Month, y=Temp)) + 
  geom_point() +
  scale_x_discrete(breaks=seq(0,203,12))
```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
names(pf)
head(pf)
pf$age_with_months <- pf$age + pf$dob_month/12
```

***

### Age with Months Means

Programming Assignment
```{r Programming Assignment}
# Create a new data frame called
# pf.fc_by_age_months that contains
# the mean friend count, the median friend
# count, and the number of users in each
# group of age_with_months. The rows of the
# data framed should be arranged in increasing
# order by the age_with_months variable.

# For example, the first two rows of the resulting
# data frame would look something like...

# age_with_months  friend_count_mean	friend_count_median	n
#              13            275.0000                   275 2
#        13.25000            133.2000                   101 11

pf.fc_by_age_months <- group_by(pf, age_with_months) %>%
  summarise( friend_count_mean=mean(friend_count),
             friend_count_median=median(friend_count),
             n=n()) %>%
  arrange(age_with_months)
?arrange
```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
ggplot(subset(pf.fc_by_age_months, age_with_months<71), aes(x=age_with_months, y=friend_count_mean)) + geom_line()
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
p1 <- ggplot(subset(pf.fc_by_age, age<71), aes(x=age, y=friend_count_mean)) + geom_line() + geom_smooth()

p2 <- ggplot(subset(pf.fc_by_age_months, age_with_months<71), aes(x=age_with_months, y=friend_count_mean)) + geom_line() + geom_smooth()

p3 <- ggplot(subset(pf.fc_by_age, age<71), aes(x=round(age/5)*5, y=friend_count_mean)) + geom_line(stat='summary', fun.y=mean)

library(gridExtra)
grid.arrange(p2,p1,p3,ncol=1)
```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

