---
title: "A/B Testing -- Lesson 1"
author: "Jeff Irion"
date: "April 5, 2016"
output: 
  html_document:
    toc: true
---

```{r, include=FALSE}
rm(list=ls())
setwd("C:/Users/Jeff/udacity/AB_Testing/")
```

### Binomial Distribution

Consider a binomial distribution with mean $p$.  If we sample that distribution $N$ times, the mean will be $p$ and the standard deviation will be $\sqrt{p(1-p)/N}$.  



### Calculating a Confidence Interval

```{r}
# total users
N <- 2000

# clicks
x <- 300

# mean
phat <- x/N

# standard deviation
s <- sqrt(phat*(1-phat)/N)

# 99% confidence level ==> use p=0.995
z <- qnorm(0.995)
```

The 99% confidence interval is (`r I(phat-s*z)`, `r I(phat+s*z)`).  



### Calculating Number of Page Views Needed

```{r}
N <- 1000
x <- 100
dmin <- 0.02
alpha <- 0.05
beta <- 0.2

# from http://www.evanmiller.org/ab-testing/sample-size.html
num_subjects <- function(alpha, beta, p, delta){
  z_alpha = qnorm(1-alpha/2)
  z_beta = qnorm(1-beta)
  
  sd1 = sqrt(2*p*(1-p))
  sd2 = sqrt(p*(1-p)+(p+delta)*(1-p-delta))
  
  return( (z_alpha * sd1 + z_beta * sd2)*(z_alpha * sd1 + z_beta * sd2)/delta^2 )
}
```

Number of subjects needed: `r I(round(num_subjects(alpha, beta, x/N, dmin)))`



### Calculating Results

```{r}
Ncont <- 10072
xcont <- 974
Nexp <- 9886
xexp <- 1242

ppool <- (974+1242)/(10072+9886)
SEpool <- sqrt(ppool*(1-ppool)*(1/Ncont+1/Nexp))

dmin <- 0.02
alpha <- 0.05

# estimated difference
dhat <- xexp/Nexp-xcont/Ncont

# margin of error
m <- qnorm(.975)*SEpool
```

$\hat d$ = `r I(dhat)`

$m$ = `r I(m)`

Confidence interval = (`r I(dhat-m)`, `r I(dhat+m)`)